---
title: "income"
author: "ZHE ZHENG"
date: "6/10/2021"
output: html_document
---
## load required packages
```{r}
library(readr)
```

## access median income file and combine it with respiratory hospitalization dataset
```{r}
# read in median income file
medianincome <- read_csv("~/Box/aim2/data/medianincome.csv", 
     col_types = cols(DP03_0062E = col_number(), 
        X4 = col_skip(), X5 = col_skip(), 
         X6 = col_skip(), X7 = col_skip(), 
         X8 = col_skip(), `zip code tabulation area` = col_character()))

medianincome <- medianincome[medianincome$DP03_0062E>10000,] # get rid of 9999 and others codes which indicate income not known
medianincome <- medianincome[complete.cases(medianincome$DP03_0062E),]# get rid of NA

colnames(medianincome)[3] <- "zip" # change the column names for identification

medianincome$zip <-formatC(as.integer(medianincome$zip),width = 5,flag = "0") # add 0 for zip codes which start with 0 (because csv file neglect the starting 0)

quantile(medianincome$DP03_0062E[medianincome$zip%in%NJresp.1$zip|medianincome$zip%in%NYresp.1$zip|medianincome$zip%in%WAresp.1$zip],c(1/3,2/3)) # get income tertile of income

# get rid of the cases that reside outside of the states
NJresp.1 <- NJresp.1[NJresp.1$zip%in%njpop$zip,] 
NYresp.1 <- NYresp.1[NYresp.1$zip%in%nypop$zip,]
WAresp.1 <- WAresp.1[WAresp.1$zip%in%WAarea$zip,]
```

# seperate the patients based on the tertile of income in the location they reside
```{r}
library(data.table )
NJincome <- merge(medianincome,NJresp.1,by.x="zip",by.y="zip")
NJincome <- data.table(NJincome)
NJincome <- NJincome[,list(resp = sum(resp),rsv=sum(rsv),bronchio=sum(bronchio),flu=sum(flu),pneusept=sum(pneumosept)), by = 'agecat1,amonthdate,DP03_0062E,zip']
resp.group1 <- NJincome[NJincome$DP03_0062E<44447,]
resp.group1 <- resp.group1[order(resp.group1$amonthdate),]
resp.group2 <- NJincome[NJincome$DP03_0062E>=44447&NJincome$DP03_0062E<62005 ,]
resp.group2 <- resp.group2[order(resp.group2$amonthdate),]
resp.group3 <- NJincome[NJincome$DP03_0062E>=62005,]
resp.group3 <- resp.group3[order(resp.group3$amonthdate),]

resp.group_nj <- list()
resp.group_nj <- list(resp.group1,resp.group2,resp.group3)
```


```{r}
NYincome <- merge(medianincome,NYresp.1,by.x="zip",by.y="zip")
NYincome <- data.table(NYincome)
NYincome <- NYincome[,list(resp = sum(resp),rsv=sum(rsv),bronchio=sum(bronchio),flu=sum(flu),pneusept=sum(pneumosept)), by = 'agecat1,amonthdate,DP03_0062E,zip']
resp.group1 <- NYincome[NYincome$DP03_0062E<44447,]
resp.group1 <- resp.group1[order(resp.group1$amonthdate),]
resp.group2 <- NYincome[NYincome$DP03_0062E>=44447&NYincome$DP03_0062E<62005 ,]
resp.group2 <- resp.group2[order(resp.group2$amonthdate),]
resp.group3 <- NYincome[NYincome$DP03_0062E>=62005 ,]
resp.group3 <- resp.group3[order(resp.group3$amonthdate),]

resp.group_ny <- list()
resp.group_ny <- list(resp.group1,resp.group2,resp.group3)

```

```{r}
WAincome <-   merge(medianincome,WAresp.1,by.x="zip",by.y="zip")
WAincome <- data.table(WAincome)
WAincome <- WAincome[,list(resp = sum(resp),rsv=sum(rsv),bronchio=sum(bronchio),flu=sum(flu),pneusept=sum(pneumosept)), by = 'agecat1,amonthdate,DP03_0062E,zip']
resp.group1 <- WAincome[WAincome$DP03_0062E<44447,]
resp.group1 <- resp.group1[order(resp.group1$amonthdate),]
resp.group2 <- WAincome[WAincome$DP03_0062E>=44447&WAincome$DP03_0062E<62005,]
resp.group2 <- resp.group2[order(resp.group2$amonthdate),]
resp.group3 <- WAincome[WAincome$DP03_0062E>=62005,]
resp.group3 <- resp.group3[order(resp.group3$amonthdate),]

resp.group_wa <- list()
resp.group_wa <- list(resp.group1,resp.group2,resp.group3)
```

# create the response variable array
# the first dimension is respiratory hospitalizations by month
# the second dimension is respiratory hospitalizations by age group
# the third dimension is respiratory hospitalizations by income group
# the forth dimension is respiratory hospitalizations by state
```{r}
y_income <- array(data=NA,dim = c(108,9,3,3))
colnames(y) <- c("<1","1-<2","2-<5","5-9","10-19","20-44","45-64","65-84",">85")

for (i in 1:3) {
  
rsv.infant <- resp.group_nj[[i]][agecat1==0,]
rsv.infant <- rsv.infant[,list(resp = sum(resp)),by = 'amonthdate']
rsv.infant <- rsv.infant[order(amonthdate),]
colnames(rsv.infant)[2] <- "infant"

rsv.12to24month <- resp.group_nj[[i]][agecat1==1,]
rsv.12to24month <- rsv.12to24month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.12to24month <- rsv.12to24month[order(amonthdate),]
colnames(rsv.12to24month)[2] <- "oney"

rsv.25to60month <- resp.group_nj[[i]][1<agecat1&agecat1<5,]
rsv.25to60month <- rsv.25to60month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.25to60month <- rsv.25to60month[order(amonthdate),]
colnames(rsv.25to60month)[2] <- "twoy"

rsv.5to9 <- resp.group_nj[[i]][agecat1==5,]
rsv.5to9 <- rsv.5to9[,list(resp = sum(resp)),by = 'amonthdate']
rsv.5to9 <-rsv.5to9[order(amonthdate),]
colnames(rsv.5to9)[2] <- "Five"

rsv.10to19 <- resp.group_nj[[i]][5<agecat1&agecat1<8,]
rsv.10to19 <- rsv.10to19[,list(resp = sum(resp)),by = 'amonthdate']
rsv.10to19 <-rsv.10to19[order(amonthdate),]
colnames(rsv.10to19)[2] <- "Ten"

rsv.20to44 <- resp.group_nj[[i]][7<agecat1&agecat1<13,]
rsv.20to44 <- rsv.20to44[,list(resp = sum(resp)),by = 'amonthdate']
rsv.20to44 <-rsv.20to44[order(amonthdate),]
colnames(rsv.20to44)[2] <- "Twenty"

rsv.45to64 <- resp.group_nj[[i]][12<agecat1&agecat1<17,]
rsv.45to64 <- rsv.45to64[,list(resp = sum(resp)),by = 'amonthdate']
rsv.45to64 <-rsv.45to64[order(amonthdate),]
colnames(rsv.45to64)[2] <- "Fourtyfive"

rsv.65to79  <- resp.group_nj[[i]][17<=agecat1&agecat1<=20,]
rsv.65to79  <- rsv.65to79 [,list(resp = sum(resp)),by = 'amonthdate']
rsv.65to79  <- rsv.65to79 [order(amonthdate),]
colnames(rsv.65to79)[2] <- "Sixtyfive"

rsv.elderly <- resp.group_nj[[i]][21<=agecat1,]
rsv.elderly <- rsv.elderly[,list(resp = sum(resp)),by = 'amonthdate']
rsv.elderly <-rsv.elderly[order(amonthdate),]
colnames(rsv.elderly)[2] <- "Elderly"

respcount <- merge(merge(merge(merge(merge(merge(merge(merge(merge(rsv_non,
  rsv.infant,all=T),
  rsv.12to24month, all = TRUE),
  rsv.25to60month, all = TRUE),
  rsv.5to9, all = TRUE),
  rsv.10to19, all = TRUE),
  rsv.20to44,all=T),
  rsv.45to64,all=T),
  rsv.65to79,all=T),
  rsv.elderly,all=T)

respcount[is.na(respcount)]=0
respcount <- respcount[order(amonthdate),]


y_income[,1,i,1] <- respcount$infant
y_income[,2,i,1] <- respcount$oney
y_income[,3,i,1] <- respcount$twoy
y_income[,4,i,1] <- respcount$Five
y_income[,5,i,1] <- respcount$Ten
y_income[,6,i,1] <- respcount$Twenty
y_income[,7,i,1] <- respcount$Fourtyfive
y_income[,8,i,1] <- respcount$Sixtyfive
y_income[,9,i,1] <- respcount$Elderly
print(i)
}

for (i in 1:3) {
  
rsv.infant <- resp.group_ny[[i]][agecat1==0,]
rsv.infant <- rsv.infant[,list(resp = sum(resp)),by = 'amonthdate']
rsv.infant <- rsv.infant[order(amonthdate),]
colnames(rsv.infant)[2] <- "infant"

rsv.12to24month <- resp.group_ny[[i]][agecat1==1,]
rsv.12to24month <- rsv.12to24month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.12to24month <- rsv.12to24month[order(amonthdate),]
colnames(rsv.12to24month)[2] <- "oney"

rsv.25to60month <- resp.group_ny[[i]][1<agecat1&agecat1<5,]
rsv.25to60month <- rsv.25to60month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.25to60month <- rsv.25to60month[order(amonthdate),]
colnames(rsv.25to60month)[2] <- "twoy"

rsv.5to9 <- resp.group_ny[[i]][agecat1==5,]
rsv.5to9 <- rsv.5to9[,list(resp = sum(resp)),by = 'amonthdate']
rsv.5to9 <-rsv.5to9[order(amonthdate),]
colnames(rsv.5to9)[2] <- "Five"

rsv.10to19 <- resp.group_ny[[i]][5<agecat1&agecat1<8,]
rsv.10to19 <- rsv.10to19[,list(resp = sum(resp)),by = 'amonthdate']
rsv.10to19 <-rsv.10to19[order(amonthdate),]
colnames(rsv.10to19)[2] <- "Ten"

rsv.20to44 <- resp.group_ny[[i]][7<agecat1&agecat1<13,]
rsv.20to44 <- rsv.20to44[,list(resp = sum(resp)),by = 'amonthdate']
rsv.20to44 <-rsv.20to44[order(amonthdate),]
colnames(rsv.20to44)[2] <- "Twenty"

rsv.45to64 <- resp.group_ny[[i]][12<agecat1&agecat1<17,]
rsv.45to64 <- rsv.45to64[,list(resp = sum(resp)),by = 'amonthdate']
rsv.45to64 <-rsv.45to64[order(amonthdate),]
colnames(rsv.45to64)[2] <- "Fourtyfive"

rsv.65to79  <- resp.group_ny[[i]][17<=agecat1&agecat1<=20,]
rsv.65to79  <- rsv.65to79 [,list(resp = sum(resp)),by = 'amonthdate']
rsv.65to79  <- rsv.65to79 [order(amonthdate),]
colnames(rsv.65to79)[2] <- "Sixtyfive"

rsv.elderly <- resp.group_ny[[i]][21<=agecat1,]
rsv.elderly <- rsv.elderly[,list(resp = sum(resp)),by = 'amonthdate']
rsv.elderly <-rsv.elderly[order(amonthdate),]
colnames(rsv.elderly)[2] <- "Elderly"

respcount <- merge(merge(merge(merge(merge(merge(merge(merge(
  rsv.infant,
  rsv.12to24month, all = TRUE),
  rsv.25to60month, all = TRUE),
  rsv.5to9, all = TRUE),
  rsv.10to19, all = TRUE),
  rsv.20to44,all=T),
  rsv.45to64,all=T),
  rsv.65to79,all=T),
  rsv.elderly,all=T) 

respcount[is.na(respcount)]=0
respcount <- respcount[order(amonthdate),]


y_income[,1,i,2] <- respcount$infant
y_income[,2,i,2] <- respcount$oney
y_income[,3,i,2] <- respcount$twoy
y_income[,4,i,2] <- respcount$Five
y_income[,5,i,2] <- respcount$Ten
y_income[,6,i,2] <- respcount$Twenty
y_income[,7,i,2] <- respcount$Fourtyfive
y_income[,8,i,2] <- respcount$Sixtyfive
y_income[,9,i,2] <- respcount$Elderly
print(i)
}

for (i in 1:3) {
  
rsv.infant <- resp.group_wa[[i]][agecat1==0,]
rsv.infant <- rsv.infant[,list(resp = sum(resp)),by = 'amonthdate']
rsv.infant <- rsv.infant[order(amonthdate),]
colnames(rsv.infant)[2] <- "infant"

rsv.12to24month <- resp.group_wa[[i]][agecat1==1,]
rsv.12to24month <- rsv.12to24month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.12to24month <- rsv.12to24month[order(amonthdate),]
colnames(rsv.12to24month)[2] <- "oney"

rsv.25to60month <- resp.group_wa[[i]][1<agecat1&agecat1<5,]
rsv.25to60month <- rsv.25to60month[,list(resp = sum(resp)),by = 'amonthdate']
rsv.25to60month <- rsv.25to60month[order(amonthdate),]
colnames(rsv.25to60month)[2] <- "twoy"

rsv.5to9 <- resp.group_wa[[i]][agecat1==5,]
rsv.5to9 <- rsv.5to9[,list(resp = sum(resp)),by = 'amonthdate']
rsv.5to9 <-rsv.5to9[order(amonthdate),]
colnames(rsv.5to9)[2] <- "Five"

rsv.10to19 <- resp.group_wa[[i]][5<agecat1&agecat1<8,]
rsv.10to19 <- rsv.10to19[,list(resp = sum(resp)),by = 'amonthdate']
rsv.10to19 <-rsv.10to19[order(amonthdate),]
colnames(rsv.10to19)[2] <- "Ten"

rsv.20to44 <- resp.group_wa[[i]][7<agecat1&agecat1<13,]
rsv.20to44 <- rsv.20to44[,list(resp = sum(resp)),by = 'amonthdate']
rsv.20to44 <-rsv.20to44[order(amonthdate),]
colnames(rsv.20to44)[2] <- "Twenty"

rsv.45to64 <- resp.group_wa[[i]][12<agecat1&agecat1<17,]
rsv.45to64 <- rsv.45to64[,list(resp = sum(resp)),by = 'amonthdate']
rsv.45to64 <-rsv.45to64[order(amonthdate),]
colnames(rsv.45to64)[2] <- "Fourtyfive"

rsv.65to79  <- resp.group_wa[[i]][17<=agecat1&agecat1<=20,]
rsv.65to79  <- rsv.65to79 [,list(resp = sum(resp)),by = 'amonthdate']
rsv.65to79  <- rsv.65to79 [order(amonthdate),]
colnames(rsv.65to79)[2] <- "Sixtyfive"

rsv.elderly <- resp.group_wa[[i]][21<=agecat1,]
rsv.elderly <- rsv.elderly[,list(resp = sum(resp)),by = 'amonthdate']
rsv.elderly <-rsv.elderly[order(amonthdate),]
colnames(rsv.elderly)[2] <- "Elderly"

respcount <- merge(merge(merge(merge(merge(merge(merge(merge(
  rsv.infant,
  rsv.12to24month, all = TRUE),
  rsv.25to60month, all = TRUE),
  rsv.5to9, all = TRUE),
  rsv.10to19, all = TRUE),
  rsv.20to44,all=T),
  rsv.45to64,all=T),
  rsv.65to79,all=T),
  rsv.elderly,all=T) 

respcount[is.na(respcount)]=0
respcount <- respcount[order(amonthdate),]


y_income[,1,i,3] <- respcount$infant
y_income[,2,i,3] <- respcount$oney
y_income[,3,i,3] <- respcount$twoy
y_income[,4,i,3] <- respcount$Five
y_income[,5,i,3] <- respcount$Ten
y_income[,6,i,3] <- respcount$Twenty
y_income[,7,i,3] <- respcount$Fourtyfive
y_income[,8,i,3] <- respcount$Sixtyfive
y_income[,9,i,3] <- respcount$Elderly
print(i)
}

```

# create the input variable array

## RSV
## the first dimension is RSV hospitalizations in children under 2 years old by month
## the second dimension will be the same because we use RSV hospitalizations in children under 2 years old for all age group as indicator
## the third dimension is RSV hospitalizations in children under 2 years old by income group
## the forth dimension is RSV hospitalizations in children under 2 years old by state

```{r}
rsv_low_nj <- resp.group_nj[[1]][agecat1<=1,]
rsv_low_nj <-  rsv_low_nj[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_low_nj <- rsv_low_nj[order(amonthdate),]

rsv_mid_nj <- resp.group_nj[[2]][agecat1<=1,]
rsv_mid_nj <-  rsv_mid_nj[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_mid_nj <- rsv_mid_nj[order(amonthdate),]

rsv_high_nj <- resp.group_nj[[3]][agecat1<=1,]
rsv_high_nj <-  rsv_high_nj[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_high_nj <- rsv_high_nj[order(amonthdate),]

rsv_low_ny <- resp.group_ny[[1]][agecat1<=1,]
rsv_low_ny <-  rsv_low_ny[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_low_ny <- rsv_low_ny[order(amonthdate),]

rsv_mid_ny <- resp.group_ny[[2]][agecat1<=1,]
rsv_mid_ny <-  rsv_mid_ny[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_mid_ny <- rsv_mid_ny[order(amonthdate),]

rsv_high_ny <- resp.group_ny[[3]][agecat1<=1,]
rsv_high_ny <-  rsv_high_ny[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_high_ny <- rsv_high_ny[order(amonthdate),]


rsv_low_wa <- resp.group_wa[[1]][agecat1<=1,]
rsv_low_wa <-  rsv_low_wa[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_low_wa <- rsv_low_wa[order(amonthdate),]

rsv_mid_wa <- resp.group_wa[[2]][agecat1<=1,]
rsv_mid_wa <-  rsv_mid_wa[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_mid_wa <- rsv_mid_wa[order(amonthdate),]

rsv_high_wa <- resp.group_wa[[3]][agecat1<=1,]
rsv_high_wa <-  rsv_high_wa[,list(rsv=sum(rsv)), by = 'amonthdate']
rsv_high_wa <- rsv_high_wa[order(amonthdate),]

rsv_income <- array(data =c(c(rep(rsv_low_nj$rsv,9),rep(rsv_mid_nj$rsv,9),rep(rsv_high_nj$rsv,9)),c(rep(rsv_low_ny$rsv,9),rep(rsv_mid_ny$rsv,9),rep(rsv_high_ny$rsv,9)),c(rep(rsv_low_wa$rsv,9),rep(rsv_mid_wa$rsv,9),rep(rsv_high_wa$rsv,9))),dim = c(108,9,3,3))
```

## Flu
## the first dimension is flu hospitalizations in entire population by month
## the second dimension will be the same because we use flu hospitalizations in entire population for all age group as indicator
## the third dimension is flu hospitalizations in entire population by income group
## the forth dimension is flu hospitalizations in entire population by state

```{r}
flu_low_nj <- resp.group_nj[[1]][,list(flu=sum(flu)), by = 'amonthdate']
flu_low_nj <- flu_low_nj[order(amonthdate),]

flu_mid_nj <- resp.group_nj[[2]][,list(flu=sum(flu)), by = 'amonthdate']
flu_mid_nj <- flu_mid_nj[order(amonthdate),]

flu_high_nj <- resp.group_nj[[3]][,list(flu=sum(flu)), by = 'amonthdate']
flu_high_nj <- flu_high_nj[order(amonthdate),]

flu_low_ny <- resp.group_ny[[1]][,list(flu=sum(flu)), by = 'amonthdate']
flu_low_ny <- flu_low_ny[order(amonthdate),]

flu_mid_ny <- resp.group_ny[[2]][,list(flu=sum(flu)), by = 'amonthdate']
flu_mid_ny <- flu_mid_ny[order(amonthdate),]

flu_high_ny <- resp.group_ny[[3]][,list(flu=sum(flu)), by = 'amonthdate']
flu_high_ny <- flu_high_ny[order(amonthdate),]


flu_low_wa <- resp.group_wa[[1]][,list(flu=sum(flu)), by = 'amonthdate']
flu_low_wa <- flu_low_wa[order(amonthdate),]

flu_mid_wa <- resp.group_wa[[2]][,list(flu=sum(flu)), by = 'amonthdate']
flu_mid_wa <- flu_mid_wa[order(amonthdate),]

flu_high_wa <- resp.group_wa[[3]][,list(flu=sum(flu)), by = 'amonthdate']
flu_high_wa <- flu_high_wa[order(amonthdate),]

flu_income <- array(data =c(c(rep(flu_low_nj$flu,9),rep(flu_mid_nj$flu,9),rep(flu_high_nj$flu,9)),c(rep(flu_low_ny$flu,9),rep(flu_mid_ny$flu,9),rep(flu_high_ny$flu,9)),c(rep(flu_low_wa$flu,9),rep(flu_mid_wa$flu,9),rep(flu_high_wa$flu,9))),dim = c(108,9,3,3))

```

# this creates a array that neglect the state difference
```{r}
y_income_whole <- rowSums(y_income,dims = 3)
rsv_income_whole <- rowSums(rsv_income,dims = 3)
flu_income_whole <- rowSums(flu_income,dims = 3)
```