---
title: "Post_table_plot"
author: "ZHE ZHENG"
date: '2022-09-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load require packages
```{r}
library(ggplot2)
```

# get posterior samples (delete the burn-in period) and calculate respiratory hospitalizations attributable to RSV infection
```{r}
post1 <- as.data.frame(as.matrix(rsv_resp[[1]][-c(1:1750),]))
  post2 <- as.data.frame(as.matrix(rsv_resp[[2]][-c(1:1750),]))
  post3 <- as.data.frame(as.matrix(rsv_resp[[3]][-c(1:1750),]))
  post <- bind_rows(post1,post2,post3)
  lambda <- post[, grep("lambda", colnames(post), fixed=T)]
  rd2.resp <- post[, grep("rd2[", colnames(post), fixed=T)]
  
  rsv_count <- array(data = NA,dim = c(3750,108,9,3))
for (i in 1:108) {
  for (j in 1:9) {
    for (k in 1:3) {
      rsv_count[,i,j,k] <- rsv_income_whole[i,j,k]*rd2.resp[,j+9*(k-1)]
    }}}

  rsv_count_age <- array(data = NA,dim = c(3750,9,3))
  for (i in 1:3750) {    
    for (j in 1:9) {
      for (k in 1:3) {
       rsv_count_age[i,j,k] <- sum(rsv_count[i,,j,k])}}}
  
  rsv_percent <- array(data = NA,dim = c(3750,9,3))
for (j in 1:9) {
  for (k in 1:3) {
    rsv_percent[,j,k] <- as.numeric(rowSums(rsv_count[,,j,k])/rowSums(lambda[,((j-1)*108+(k-1)*108*9+1):((j-1)*108+(k-1)*108*9+108)]))
  }}

  round(apply(rsv_percent,c(2), mean)*100,1)
  round(apply(rsv_percent, c(2,3), quantile , probs = 0.025)*100,1)
  round(apply(rsv_percent, c(2,3), quantile , probs = 0.975)*100,1)
```

# get population size in each age group
```{r}
# read in zip code level population size of each age group
age_pop <- readRDS("~/Library/CloudStorage/Box-Box/aim2/aim2/age_pop.rds")

# seperate the zip code by tertile of income
age_pop_1 <- age_pop[age_pop$ZIP%in%resp.group_nj[[1]]$zip|age_pop$ZIP%in%resp.group_ny[[1]]$zip|age_pop$ZIP%in%resp.group_wa[[1]]$zip,]
age_pop_2 <- age_pop[age_pop$ZIP%in%resp.group_nj[[2]]$zip|age_pop$ZIP%in%resp.group_ny[[2]]$zip|age_pop$ZIP%in%resp.group_wa[[2]]$zip,]
age_pop_3 <- age_pop[age_pop$ZIP%in%resp.group_nj[[3]]$zip|age_pop$ZIP%in%resp.group_ny[[3]]$zip|age_pop$ZIP%in%resp.group_wa[[3]]$zip,]

# create a data frame that contains population by age (row) and by income (columns)
age_number <- matrix(data=NA,nrow=9,ncol = 3)
colnames(age_number) <- c("High","Medium","Low")
rownames(age_number) <- c("<1","1","2-4","5-9","10-19","20-44","45-64","65-84","85+")
age_number[1,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[2,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[3,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..Under.5.years)/5*3
age_number[4,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..5.to.9.years)
age_number[5,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..10.to.14.years,age_pop_1$Estimate..SEX.AND.AGE..15.to.19.years)
age_number[6,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..20.to.24.years,age_pop_1$Estimate..SEX.AND.AGE..25.to.34.years,age_pop_1$Estimate..SEX.AND.AGE..35.to.44.years)
age_number[7,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..45.to.54.years,age_pop_1$Estimate..SEX.AND.AGE..55.to.59.years,age_pop_1$Estimate..SEX.AND.AGE..60.to.64.years)
age_number[8,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..65.to.74.years,
                       age_pop_1$Estimate..SEX.AND.AGE..75.to.84.years)
age_number[9,1] <- sum(age_pop_1$Estimate..SEX.AND.AGE..85.years.and.over)


age_number[1,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[2,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[3,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..Under.5.years)/5*3
age_number[4,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..5.to.9.years)
age_number[5,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..10.to.14.years,age_pop_2$Estimate..SEX.AND.AGE..15.to.19.years)
age_number[6,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..20.to.24.years,age_pop_2$Estimate..SEX.AND.AGE..25.to.34.years,age_pop_2$Estimate..SEX.AND.AGE..35.to.44.years)
age_number[7,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..45.to.54.years,age_pop_2$Estimate..SEX.AND.AGE..55.to.59.years,age_pop_2$Estimate..SEX.AND.AGE..60.to.64.years)
age_number[8,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..65.to.74.years,
                       age_pop_2$Estimate..SEX.AND.AGE..75.to.84.years)
age_number[9,2] <- sum(age_pop_2$Estimate..SEX.AND.AGE..85.years.and.over)



age_number[1,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[2,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..Under.5.years)/5
age_number[3,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..Under.5.years)/5*3
age_number[4,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..5.to.9.years)
age_number[5,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..10.to.14.years,age_pop_3$Estimate..SEX.AND.AGE..15.to.19.years)
age_number[6,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..20.to.24.years,age_pop_3$Estimate..SEX.AND.AGE..25.to.34.years,age_pop_3$Estimate..SEX.AND.AGE..35.to.44.years)
age_number[7,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..45.to.54.years,age_pop_3$Estimate..SEX.AND.AGE..55.to.59.years,age_pop_3$Estimate..SEX.AND.AGE..60.to.64.years)
age_number[8,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..65.to.74.years,
                       age_pop_3$Estimate..SEX.AND.AGE..75.to.84.years)
age_number[9,3] <- sum(age_pop_3$Estimate..SEX.AND.AGE..85.years.and.over)
```

# create plot 1
```{r}
incidence <- array(data = NA,dim = c(3750,9,3))
  for (j in 1:9) {
      for (k in 1:3) {
        for (p in 1:3750) {
   incidence[p,j,k] <- (rsv_count_age[p,j,k]/age_number[j,k])/9*100000}}}

dimnames(incidence)[2] <- list(age_group=c("<1","1","2-4","5-9","10-19","20-44","45-64","65-84","85+"))

  round(apply(incidence,2, mean))
  round(apply(incidence, 2, quantile , probs = 0.025))
  round(apply(incidence,2, quantile , probs = 0.975))
  
dimnames(incidence)[3] <- list(income_group=c("Low","Medium","High"))

df0 <- as.data.frame.table(incidence)
df0 <- df0[,-1]
names(df0) <- c("age_group","income_group","RSVHosp")

df2 <- df0 %>%
  group_by(age_group,income_group) %>%
  summarise(RSVHosp_incidence = mean(RSVHosp, na.rm = TRUE),
            quan_0.025 = quantile(RSVHosp, 0.025,na.rm = TRUE),
            quan_0.975 = quantile(RSVHosp, 0.975,na.rm = TRUE),
            sd=sd(RSVHosp, na.rm = TRUE))
df3 <- df2
df3$RSVHosp_incidence <- signif(df3$RSVHosp_incidence,2)

jpeg(file = "/Users/zhezheng/Library/CloudStorage/Box-Box/aim2/draft/revision_pneu/F1_0823.jpeg",width=2500,height=1200,res=300)   
ggplot(df3, aes(x=age_group,y=RSVHosp_incidence,label=RSVHosp_incidence,color = income_group,group=income_group)) +
  geom_point(aes(x=age_group,y=RSVHosp_incidence),position=position_dodge(0.6), size = 1)+
 geom_text(position = position_dodge(width =0.7), vjust = -df2$sd/50-1, size = 3,
              fontface = "bold", show.legend = FALSE) +
  scale_y_continuous(breaks=seq(0, 3000, 250), limits=c(0, 3100))+
  geom_errorbar(
    aes(ymin = quan_0.025, ymax = quan_0.975,color = income_group),
    position = position_dodge(0.6),width=.1
    )+
  scale_color_manual(values = c("#00AFBB", "#E7B800","#FC4E07"))+
  labs(y="Estimated RSV hospitalizations\nper 100,000 people per year", x="Age group",color="Income level",size = 8)+
  theme_classic()+
  theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),
        legend.title = element_text(size=8),
        legend.text = element_text(size = 8))
dev.off()

```

# get reported RSV by age and income group
```{r}
rsv_report <- matrix(data=NA,nrow=9,ncol=3)
rownames(rsv_report) <- c("<1","1-<2","2-<5","5-9","10-19","20-44","45-64","65-84",">85")

for (i in 1:3) {
  
rsv_report[1,i] <- sum(resp.group_nj[[i]][agecat1==0,][,6],resp.group_ny[[i]][agecat1==0,][,6],resp.group_wa[[i]][agecat1==0,][,6])

rsv_report[2,i] <- sum(resp.group_nj[[i]][agecat1==1,][,6],resp.group_ny[[i]][agecat1==1,][,6],resp.group_wa[[i]][agecat1==1,][,6])

rsv_report[3,i]<- sum(resp.group_nj[[i]][1<agecat1&agecat1<5,][,6],resp.group_ny[[i]][1<agecat1&agecat1<5,][,6],resp.group_wa[[i]][1<agecat1&agecat1<5,][,6])

rsv_report[4,i] <- sum(resp.group_nj[[i]][agecat1==5,][,6],resp.group_ny[[i]][agecat1==5,][,6],resp.group_wa[[i]][agecat1==5,][,6])

rsv_report[5,i] <- sum(resp.group_nj[[i]][5<agecat1&agecat1<8,][,6],resp.group_ny[[i]][5<agecat1&agecat1<8,][,6],resp.group_wa[[i]][5<agecat1&agecat1<8,][,6])

rsv_report[6,i]<- sum(resp.group_nj[[i]][7<agecat1&agecat1<13,][,6],resp.group_ny[[i]][7<agecat1&agecat1<13,][,6],resp.group_wa[[i]][7<agecat1&agecat1<13,][,6])

rsv_report[7,i] <- sum(resp.group_nj[[i]][12<agecat1&agecat1<17,][,6],resp.group_ny[[i]][12<agecat1&agecat1<17,][,6],resp.group_wa[[i]][12<agecat1&agecat1<17,][,6])

rsv_report[8,i] <- sum(resp.group_nj[[i]][17<=agecat1&agecat1<=20,][,6],resp.group_ny[[i]][17<=agecat1&agecat1<=20,][,6],resp.group_wa[[i]][17<=agecat1&agecat1<=20,][,6])

rsv_report[9,i] <- sum(resp.group_nj[[i]][21<=agecat1,][,6],resp.group_ny[[i]][21<=agecat1,][,6],resp.group_wa[[i]][21<=agecat1,][,6])

print(i)
}

```

# create figure 2. The reporting ratio of RSV hospitalizations across age and income groups
```{r}
# calculate the report ratio across all posterior samples
# need to divide first and then calculate the interval
rsv_ratio <- array(data = NA,dim = c(3750,9,3))
  for (j in 1:9) {
      for (k in 1:3) {
        for (p in 1:3750) {
   rsv_ratio[p,j,k] <- rsv_report[j,k]/rsv_count_age[p,j,k]}}}

dimnames(rsv_ratio)[2] <- list(age_group=c("<1","1","2-4","5-9","10-19","20-44","45-64","65-84","85+"))

# get a general idea of the mean and 95% CIs of the report ratio across age groups
  round(apply(rsv_ratio,2, mean)*100,2)
  round(apply(rsv_ratio, 2, quantile , probs = 0.025)*100,2)
  round(apply(rsv_ratio,2, quantile , probs = 0.975)*100,2)
  
dimnames(rsv_ratio)[3] <- list(income_group=c("Low","Medium","High"))

rsv_ratio <- as.data.frame.table(rsv_ratio)
rsv_ratio <- rsv_ratio[,-1]
names(rsv_ratio) <- c("age_group","income_group","RSVratio")

long_rsv_ratio <- rsv_ratio %>%
  group_by(age_group,income_group) %>%
  summarise(RSV_ratio = mean(RSVratio, na.rm = TRUE)*100,
            quan_0.025 = quantile(RSVratio, 0.025,na.rm = TRUE)*100,
            quan_0.975 = quantile(RSVratio, 0.975,na.rm = TRUE)*100)

long_rsv_ratio$RSV_ratio <- round(long_rsv_ratio$RSV_ratio)
long_rsv_ratio$Age <- factor(long_rsv_ratio$age_group,level=c("85+", "65-84", "45-64","20-44","10-19" , "5-9" ,"2-4", "1","<1"  ))
long_rsv_ratio$`Income group` <- factor(long_rsv_ratio$income_group,levels = c("Low","Medium","High"))
names(long_rsv_ratio)[3] <- "Report ratio"

jpeg(file = "/Users/zhezheng/Library/CloudStorage/Box-Box/aim2/draft/revision_pneu/F2_0823.jpeg",width=1600,height=1600,res=380)
ggplot(data=long_rsv_ratio,aes(x=`Income group`, y=Age)) +
    geom_point(aes(size=quan_0.975, color=`Income group`),shape=16,alpha=0.3) +
  geom_point(aes(size = `Report ratio`,colour = `Income group`),shape = 1,stroke = 0.1) +
  geom_point(aes(size=quan_0.025),fill="white",color="white") +
  scale_color_manual(values = c( "#FC4E07","#00AFBB", "#E7B800","white"),name="Income Group") +
    geom_text(aes(label=paste0(`Report ratio`,"%")), size=3, nudge_x=+0.45, nudge_y=0.00)+
    guides(size = guide_legend(override.aes = list(shape=16),order=1),
           fill = guide_legend(override.aes = list(shape=21,size=3),order=2))+
    scale_size(range = c(0.05, 14), breaks = c(10,30,50), name="Recorded Ratio",labels = paste0(c(10,30,50),"%"))+
    theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10),
          axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),
          legend.title = element_text(size=8),
          legend.text = element_text(size = 8),
          panel.background = element_blank())
dev.off()
```

## Figure 3. stacked plot for children under one year old
```{r}
library(dplyr)
library(tidyr)
rsv_count_estimate <- array(data = NA,dim = c(3750,108))
  for (i in 1:108) {
    for (j in 1:3750) {
        rsv_count_estimate[j,i] <- sum(rsv_count[j,i,1,])
    }}
rsv_count_estimate <- colMeans(rsv_count_estimate)
rsv_count_estimate_plot <- data.frame(state=rep("Whole",108),type=rep("Estimates",108),date=rsv_urban$amonthdate,mean=rsv_count_estimate/(sum(age_number[1,]))*100000)

infant.report <- bind_rows(NJresp.1[AGE<1,list(rsv=sum(rsv)),by="amonthdate"],NYresp.1[AGE<1,list(rsv=sum(rsv)),by="amonthdate"],WAresp.1[AGE<1,list(rsv=sum(rsv)),by="amonthdate"])
infant.report <-aggregate(infant.report$rsv, by=list(amonthdate=infant.report$amonthdate), FUN=sum)
infant.report <- infant.report[order(infant.report$amonthdate),]
rsv_count_real_plot <- data.frame(state=rep("Whole",108),type=rep("Report",108),date=infant.report$amonthdate,mean=infant.report$x/(sum(age_number[1,]))*100000)

infant.total <- bind_rows(NJresp.1[AGE<1,list(resp=sum(resp)),by="amonthdate"],NYresp.1[AGE<1,list(resp=sum(resp)),by="amonthdate"],WAresp.1[AGE<1,list(resp=sum(resp)),by="amonthdate"])
infant.total <-aggregate(infant.total$resp, by=list(amonthdate=infant.total$amonthdate), FUN=sum)
infant.total <- infant.total[order(infant.total$amonthdate),]
rsv_count_total_plot <- data.frame(state=rep("Whole",108),type=rep("Total",108),date=infant.total$amonthdate,mean=infant.total$x/(sum(age_number[1,]))*100000)

rsv_wide <- bind_rows(rsv_count_total_plot,rsv_count_estimate_plot,rsv_count_real_plot)
rsv_wide <- spread(rsv_wide, type, mean)

jpeg(file = "/Users/zhezheng/Library/CloudStorage/Box-Box/aim2/draft/revision_pneu/F3_0823.jpeg",width=2500,height=1200,res=300)
ggplot(data=rsv_wide, aes(x=date)) +
 geom_area(aes(y=Total,fill="Total")) +
  geom_area(aes(y=Estimates,fill="Estimated RSV")) +
    geom_area(aes(y=Report,fill="Recorded RSV"))+
labs(y=" Hospitalizations per 100,000 children under One", x="Date")+
  theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),
        legend.title = element_text(size=10),
        legend.text = element_text(size = 10),
    panel.background = element_blank())+ 
    scale_fill_manual(name="Respiratory\nHospitalizations",values=c("Total"="darkturquoise","Estimated RSV"="gold","Recorded RSV"="red"),breaks=c("Total",
                           "Estimated RSV",
                           "Recorded RSV"))
dev.off()

```

## Figure 4. stacked plot for older adults over 65 years old

```{r}
# first run a zoom function to create the zoom in for figure 4
library(ggplot)
library(ggforce)
library(grid)

# define facet_zoom2 function to use FacetZoom2 instead of FacetZoom
# (everything else is the same as facet_zoom)
facet_zoom2 <- function(x, y, xy, zoom.data, xlim = NULL, ylim = NULL, 
                        split = FALSE, horizontal = TRUE, zoom.size = 2, 
                        show.area = TRUE, shrink = TRUE) {
  x <- if (missing(x)) if (missing(xy)) NULL else lazyeval::lazy(xy) else lazyeval::lazy(x)
  y <- if (missing(y)) if (missing(xy)) NULL else lazyeval::lazy(xy) else lazyeval::lazy(y)
  zoom.data <- if (missing(zoom.data)) NULL else lazyeval::lazy(zoom.data)
  if (is.null(x) && is.null(y) && is.null(xlim) && is.null(ylim)) {
    stop("Either x- or y-zoom must be given", call. = FALSE)
  }
  if (!is.null(xlim)) x <- NULL
  if (!is.null(ylim)) y <- NULL
  ggproto(NULL, FacetZoom2,
          shrink = shrink,
          params = list(
            x = x, y = y, xlim = xlim, ylim = ylim, split = split, zoom.data = zoom.data,
            zoom.size = zoom.size, show.area = show.area,
            horizontal = horizontal
          )
  )
}
```

```{r}
rsv_count_estimate <- array(data = NA,dim = c(3750,108))
  for (i in 1:108) {
    for (j in 1:3750) {
        rsv_count_estimate[j,i] <- sum(rsv_count[j,i,c(8:9),])
    }}

rsv_count_estimate <- colMeans(rsv_count_estimate)
rsv_count_estimate_plot_old <- data.frame(state=rep("Whole",108),type=rep("Estimates",108),date=rsv_urban$amonthdate,mean=rsv_count_estimate/( 4565877)*100000)

report_old <- bind_rows(NJresp.1[65<=AGE,list(rsv=sum(rsv)),by="amonthdate"],NYresp.1[65<=AGE,list(rsv=sum(rsv)),by="amonthdate"],WAresp.1[65<=AGE,list(rsv=sum(rsv)),by="amonthdate"])
report_old  <-aggregate(report_old $rsv, by=list(amonthdate=report_old$amonthdate), FUN=sum)
report_old  <- report_old [order(report_old $amonthdate),]
rsv_count_real_plot_old <- data.frame(state=rep("Whole",108),type=rep("Report",108),date=report_old $amonthdate,mean=report_old $x/( 4565877)*100000)

total_old <- bind_rows(NJresp.1[65<=AGE,list(resp=sum(resp)),by="amonthdate"],NYresp.1[65<=AGE,list(resp=sum(resp)),by="amonthdate"],WAresp.1[65<=AGE,list(resp=sum(resp)),by="amonthdate"])
total_old  <-aggregate(total_old$resp, by=list(amonthdate=total_old$amonthdate), FUN=sum)
total_old  <- total_old [order(total_old$amonthdate),]
rsv_count_total_plot_old <- data.frame(state=rep("Whole",108),type=rep("Total",108),date=total_old$amonthdate,mean=total_old$x/( 4565877)*100000)

rsv_old <- bind_rows(rsv_count_total_plot_old,rsv_count_estimate_plot_old,rsv_count_real_plot_old)
rsv_old <- spread(rsv_old, type, mean)

##RSV respiratory hospitalizations in the elderly (65+) in New York
jpeg(file = "/Users/zhezheng/Library/CloudStorage/Box-Box/aim2/draft/revision_pneu/F4_0823.jpeg",width=2500,height=1200,res=300)
ggplot(data=rsv_old, aes(x=date)) +
 geom_area(aes(y=Total,fill="Total")) +
  geom_area(aes(y=Estimates,fill="Estimated RSV")) +
    geom_area(aes(y=Report,fill="Recorded RSV"))+
  labs(y="Hospitalizations per 100,000 people", x="Date")+
  theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),
        legend.title = element_text(size=10),
        legend.text = element_text(size = 10),
    panel.background = element_blank(),
     plot.title = element_text( face = "bold",hjust = 0.5))+ 
    scale_fill_manual(name="Respiratory\nHospitalizations",values=c("Total"="darkturquoise","Estimated RSV"="gold","Recorded RSV"="red"),breaks=c("Total",
                           "Estimated RSV",
                           "Recorded RSV"))+
  facet_zoom2(ylim = c(0,75))
dev.off()
```
